---
title: "Rethinking case attraction in Anceint Greek Infinitive Clauses"
author: "Caio B. A. Geraldes"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(rethinking)
library(dagitty)
library(ggdag)
source("./fix_pairs_rethinking.R")
```

# Agreement framework from Corbett (2006)


```{r echo=FALSE}
dagitty("dag{
 C -> A <- D
 T -> A
}") %>%
  tidy_dagitty() %>%
  ggdag(text_size = 10) +
  theme_dag_blank()
```

# Data


```{r}
source("./pred_a_info.R")
(
  dataset <- read_csv("./data.csv", show_col_types = FALSE) %>%
    mutate(pos_pred_a = factor(pred_a_pos, levels = c("A", "N", "P")),
      pred_a = pred_a
    )
)
```

# Effect of Part of Speech

```{r echo=FALSE}
dag_a <- dagitty::dagitty("
  dag{
    A
    PoS
    Cop
    D
    D -> A <- PoS <- Cop
    Cop -> A
  }
") %>%
  tidy_dagitty() %>%
  dag_label(
    labels = c(
      "A" = "outcome",
      "D" = "unobserved",
      "Cop" = "observed",
      "PoS" = "observed"
    )
  )
```

```{r echo=FALSE}
dag_a %>%
  ggdag() +
  geom_dag_node(aes(color = label)) +
  geom_dag_text() +
  theme_dag_blank()
```

# Effect of Thematic Role and Copula

```{r echo=FALSE}
dag_b <- dagitty::dagitty("
  dag{
    A
    PoS
    Cop
    D
    C -> A
    D -> A <- PoS <- Cop
    Cop -> A
    Cop -> Th -> A
  }
") %>%
  tidy_dagitty() %>%
  dag_label(
    labels = c(
      "A" = "outcome", "D" = "unobserved", "C" = "unobserved",
      "Cop" = "observed", "PoS" = "observed", "Th" = "observed"
    )
  )
dag_b %>%
  ggdag() +
  geom_dag_node(aes(color = label)) +
  geom_dag_text() +
  theme_dag_blank()
```

# Effect of Authorship, Dialect and Genre

```{r echo=FALSE}
dag_c <- dagitty::dagitty("
  dag{
    A
    PoS
    Cop
    D
    C -> A
    D -> A <- PoS <- Cop
    Cop -> A
    Cop -> Th -> A
    Auth -> Genre -> Th
    Auth -> A <- Dia <- Auth
  }
") %>%
  tidy_dagitty() %>%
  dag_label(
    labels = c(
      "A" = "outcome", "D" = "unobserved", "C" = "unobserved",
      "Cop" = "observed", "PoS" = "observed", "WO" = "observed",
      "Th" = "observed", "Auth" = "observed", "Genre" = "observed",
      "Dia" = "observed"
    )
  )


dag_c %>%
  ggdag() +
  geom_dag_node(aes(color = label)) +
  geom_dag_text() +
  theme_dag_blank()
```

# Effect of Word Order and Distance

```{r echo=FALSE}
dag_d <- dagitty::dagitty("
  dag{
    A
    PoS
    Cop
    D
    C -> A
    D -> A <- PoS <- Cop
    Cop -> A
    Cop -> Th -> A
    T -> A
    Auth -> Genre -> Th
    Auth -> A <- Dia <- Auth
    T <- WO <- Dist -> A
  }
") %>%
  tidy_dagitty() %>%
  dag_label(
    labels = c(
      "A" = "outcome", "D" = "unobserved", "T" = "unobserved",
      "C" = "unobserved", "Cop" = "observed", "PoS" = "observed",
      "WO" = "observed", "Th" = "observed", "Auth" = "observed",
      "Genre" = "observed", "Dia" = "observed", "Dist" = "observed"
    )
  )


dag_d %>%
  ggdag() +
  geom_dag_node(aes(color = label)) +
  geom_dag_text() +
  theme_dag_blank()
```

# All exposures

## Adjustment set

```{r}
dagitty::dagitty("
  dag{
    A
    PoS
    Cop
    D[latent]
    T[latent]
    C[latent]
    C -> A
    D -> A <- PoS <- Cop
    Cop -> A
    Cop -> Th -> A
    T -> A
    Auth -> Genre -> Th
    Auth -> A <- Dia <- Auth
    T <- WO <- Dist -> A
  }
") %>%
  adjustmentSets(
    exposure = c("PoS", "Cop", "Th", "Auth", "Dia", "Dist"),
    outcome = "A", effect = "direct"
  )
```

Full causal model

```{r echo=FALSE}
dag_e <- dagitty::dagitty("
  dag{
    A
    PoS
    Cop
    D
    C -> A
    D -> A <- PoS <- Cop
    Cop -> A
    Cop -> Th -> A
    T -> A
    Auth -> Genre -> Th
    Auth -> A <- Dia <- Auth
    T <- WO <- Dist -> A
  }
") %>%
  tidy_dagitty() %>%
  dag_label(
    labels = c(
      "A" = "outcome", "D" = "unobserved", "T" = "unobserved",
      "C" = "unobserved", "Cop" = "exposure", "PoS" = "exposure",
      "WO" = "adjusted", "Th" = "exposure", "Auth" = "exposure",
      "Dia" = "exposure", "Dist" = "exposure"
    )
  )

dag_e %>%
  ggdag() +
  geom_dag_node(aes(color = label)) +
  geom_dag_text() +
  theme_dag_blank()
```

# GLM

## Preparing the data
```{r}
dataset_prep <- read_csv("./data.csv") %>%
  mutate(pos_pred_a = factor(pred_a_pos, levels = c("A", "N", "P")),
    pred_a = pred_a,
    OBJ_TH = factor(
      ifelse(
        VM_LEMMA == "δοκέω",
        "Experiencer",
        ifelse(VM_MOD, ifelse(Vinf_Cop, "Experiencer", "Agent"), "Recipient")
      ), levels = c("Recipient", "Experiencer", "Agent")
    )
  )

dat <- list(
  "A" = dataset_prep$Attr,
  "Th" = dataset_prep$OBJ_TH,
  "PoS" = dataset_prep$pos_pred_a,
  "Cop" = dataset_prep$Vinf_Cop,
  "Auth" = as.factor(dataset_prep$AUTHOR),
  "Dia" = as.factor(dataset_prep$DIALECT),
  "Prec" = (dataset_prep$DIST_OBJ_PRED < 0),
  "Dist" = abs(dataset_prep$DIST_OBJ_PRED),
  alpha_pos = rep(2, 2),
  alpha_th = rep(2, 2)
)
```

## Formula list

```{r}
flist <- alist(
  # Attraction Model
  A ~ bernoulli(p),
  logit(p) <- a +
    b_c * Cop +
    b_pos * sum(delta_j_pos[1:PoS]) +
    b_th * sum(delta_j_th[1:Th]) +
    z_auth[Auth] * s_auth +
    b_dia_bar + z_dia[Dia] * s_dia +
    b_dist * Dist +
    b_wo * Prec,
  # base
  a ~ normal(0, 1),
  # copula
  b_c ~ normal(0, 1),
  # pos
  b_pos ~ normal(0, 0.5),
  vector[3]:delta_j_pos <<- append_row(0, delta_pos),
  simplex[2]:delta_pos ~ dirichlet(alpha_pos),
  transpars > real[1]:d_adj <<- b_pos + delta_pos[1],
  transpars > real[1]:d_noun <<- b_pos + delta_pos[1] + delta_pos[2],
  # th
  b_th ~ normal(0, 0.5),
  vector[3]:delta_j_th <<- append_row(0, delta_th),
  simplex[2]:delta_th ~ dirichlet(alpha_th),
  transpars > real[1]:d_exp <<- b_th + delta_th[1],
  transpars > real[1]:d_agent <<- b_th + delta_th[1] + delta_th[2],
  # auth
  z_auth[Auth] ~ normal(0, 1),
  s_auth ~ exponential(1),
  gq > vector[Auth]:b_auth <<- z_auth * s_auth,
  # dia
  b_dia_bar ~ normal(0, 1),
  z_dia[Dia] ~ normal(0, 1),
  s_dia ~ exponential(1),
  gq > vector[Dia]:b_dia <<- b_dia_bar + z_dia * s_dia,
  # Dist
  b_dist ~ normal(0, 1),
  # WO
  b_wo ~ normal(0, 1)
)
```

## Running the model

```{r}
model <- ulam(
  flist = flist,
  data = dat,
  chains = 4, cores = 4,
  log_lik = TRUE,
  file = "./model_glm"
)
precis(model, depth = 2)
```

# Results

## Part of Speech

### Effects
```{r}
precis(model, pars = c("b_pos", "delta_pos"), depth = 2)
pairs(
  model,
  pars = c("b_pos", "delta_pos"), labels = c("Participle", "Adjective", "Noun")
)
```

### Cummulative effects

```{r}
a <- precis(model, pars = c("b_pos", "d_adj", "d_noun"), depth = 2)
tibble(
  PoS = c("Participle", "Adjective", "Noun"),
  `Cummulative effect` = a@row.names,
  mean = round(a@.Data[[1]], 2),
  sd = round(a@.Data[[2]], 2),
  `5.5%` = round(a@.Data[[3]], 2),
  `94.5%` = round(a@.Data[[4]], 2),
  rhat = round(a@.Data[[5]], 2),
  ess_bulk = round(a@.Data[[6]], 2)
)
```

## Thematic Roles

### Effects
```{r}
precis(model, pars = c("b_th", "delta_th"), depth = 2)
pairs(
  model,
  pars = c("b_th", "delta_th"), labels = c("Recipient", "Experiencer", "Agent")
)
```

### Cummulative effects

```{r}
a <- precis(model, pars = c("b_th", "d_exp", "d_agent"), depth = 2)
tibble(
  Theta = c("Recipient", "Experiencer", "Agent"),
  `Cummulative effect` = a@row.names,
  mean = round(a@.Data[[1]], 2),
  sd = round(a@.Data[[2]], 2),
  `5.5%` = round(a@.Data[[3]], 2),
  `94.5%` = round(a@.Data[[4]], 2),
  rhat = round(a@.Data[[5]], 2),
  ess_bulk = round(a@.Data[[6]], 2)
)
```


## Copula and Distance
```{r}
precis(model, pars = c("b_c", "b_dist"), depth = 2)
```

### Copula

```{r echo=FALSE}
samples <- extract.samples(model)
density_b_c <- density(samples$b_c, bw = 0.036)
density_b_c <- tibble(x = density_b_c$x, y = density_b_c$y)
boundaries <- HPDI(samples$b_c)
density_b_c %>%
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  geom_area(
    data = subset(
      density_b_c,
      x > boundaries[1] & x < boundaries[2]
    ),
    mapping = aes(x = x, y = y),
    fill = "purple",
    alpha = 0.5
  ) +
  xlab("b_cop") +
  ylab("density")
```

### Distance

```{r echo=TRUE}
density_b_dist <- density(samples$b_dist, bw = 0.0035)
density_b_dist <- tibble(x = density_b_dist$x, y = density_b_dist$y)
boundaries <- HPDI(samples$b_dist)
density_b_dist %>%
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  geom_area(
    data = subset(
      density_b_dist,
      x > boundaries[1] & x < boundaries[2]
    ),
    mapping = aes(x = x, y = y),
    fill = "purple",
    alpha = 0.5
  ) +
  xlab("b_dist") +
  ylab("density")
```

## Dialect
```{r}
precis(model, pars = c("b_dia_bar", "s_dia", "b_dia"), depth = 2)
```

```{r echo=FALSE}
density_b_attic <- density(samples$b_dia[, 1], bw = 0.06)
density_b_attic <- tibble(x = density_b_attic$x, y = density_b_attic$y)
boundaries <- HPDI(samples$b_dia[, 1])
density_b_attic %>%
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  geom_area(
    data = subset(
      density_b_attic,
      x > boundaries[1] & x < boundaries[2]
    ),
    mapping = aes(x = x, y = y),
    fill = "purple",
    alpha = 0.5
  ) +
  xlab("b_attic") +
  ylab("density")

density_b_jonic <- density(samples$b_dia[, 2], bw = 0.06)
density_b_jonic <- tibble(x = density_b_jonic$x, y = density_b_jonic$y)
boundaries <- HPDI(samples$b_dia[, 2])
density_b_jonic %>%
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  geom_area(
    data = subset(
      density_b_jonic,
      x > boundaries[1] & x < boundaries[2]
    ),
    mapping = aes(x = x, y = y),
    fill = "purple",
    alpha = 0.5
  ) +
  xlab("b_jonic") +
  ylab("density")
```

#### Contrast
```{r echo=FALSE}
density_diff_dia <- density(samples$b_dia[, 1] - samples$b_dia[, 2], bw = 0.06)
density_diff_dia <- tibble(x = density_diff_dia$x, y = density_diff_dia$y)
boundaries <- HPDI(samples$b_dia[, 1] - samples$b_dia[, 2])
density_diff_dia %>%
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  geom_area(
    data = subset(
      density_diff_dia,
      x > boundaries[1] & x < boundaries[2]
    ),
    mapping = aes(x = x, y = y),
    fill = "purple",
    alpha = 0.5
  ) +
  geom_vline(aes(xintercept = 0)) +
  xlab("b_attic - b_jonic") +
  ylab("density")
```

## Author

```{r}
precis(model, pars = c("b_auth"), depth = 2)
```

```{r}
plot(precis(model, pars = "b_auth", depth = 2),
  labels = c(
    "Aeschines",
    "Aeschylus",
    "Andocides",
    "Antiphon",
    "Aristophanes",
    "Demosthenes",
    "Euripides",
    "Herodotus",
    "Isaeus",
    "Isocrates",
    "Lycurgus",
    "Lysias",
    "Plato",
    "Sophocles",
    "Thucydides",
    "Xenophon"
  )
)
```

```{r echo=FALSE, warning=FALSE}
b_auth <- samples$b_auth
as_tibble(b_auth) %>%
  rename(
    "Aeschines" = V1,
    "Aeschylus" = V2,
    "Andocides" = V3,
    "Antiphon" = V4,
    "Aristophanes" = V5,
    "Demosthenes" = V6,
    "Euripides" = V7,
    "Herodotus" = V8,
    "Isaeus" = V9,
    "Isocrates" = V10,
    "Lycurgus" = V11,
    "Lysias" = V12,
    "Plato" = V13,
    "Sophocles" = V14,
    "Thucydides" = V15,
    "Xenophon" = V16
  ) %>%
  pivot_longer(cols = everything()
  ) %>%
  filter(value > -2, value < 2) %>%
  ggplot(aes(x = value)) +
  geom_density(aes(color = name)) +
  facet_wrap(vars(name)) +
  theme(legend.position = "none")
```
